Topic modeling on EU and the USA research projects

### 2017-06-27 thesis_report/draft_thesis.tex
Report: fixed all mistakes

\usepackage[justification=centering]{caption}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{algorithm2e}
\usepackage{url}
\usepackage{subfig}
\usepackage{listings}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{geometry}

@@

Nowadays, there are some funding opportunities for researchers, but are 
really competitive. The European Union countries are encouraged to
invest 3\% of their GDP in research and development by 2020. EU's
latest research programme, Horizon 2020, allows researchers to
identify new opportunities and directions in almost any field of
research. The programme's goals are to strengthen the EU's position in
science, support industrial innovation and address social concerns,
such as climate change, green transportation and energy.

@@

Latent Semantic Indexing was one of the earliest methods for finding
relationships between documents and the words that occur in
them\cite{Deerwester90indexingby}. In LSI a term-document matrix is
created by analyzing the corpus, where the rows correspond to words
and columns to documents. Each element in this sparse matrix describes
the number of times a word occurs in a document but this term count
can also be weighted with for instance TF-IDF\footnote{TF-IDF stands 
for term frequency - inverse document frequency, and the tf-idf weight is 
a weight often used in information retrieval and text mining. This weight 
is a statistical measure used to evaluate how important a word is to 
a document in a collection or corpus. The importance increases
proportionally to the number of times a word appears in the document (TF) but
is offset by the frequency of the word in the corpus (IDF).}. Letting
each word represent a dimension in a very high dimensional space, a
document can be seen as a vector with components corresponding to its
weighted term counts\cite{Salton:1988:TAA:54259.54260}.

@@

\subsection{EU dataset}

CORDIS (Community Research and Development Information Service) is the European
Commission's primary public repository and portal to disseminate information on
all EU-funded research projects and their results in the broadest sense. The
website includes all public information held by the Commission (project
fact sheets, publishable reports and deliverables), editorial content to support
communication and exploitation (news, events, success stories, magazines) and
comprehensive links to external sources such as open access publications and
websites. CORDIS is managed by the Publications Office of the European Union, on
behalf of the European Commission's research Directorates-General and Agencies.
CORDIS content dates back to the origin of the service in 1990 and the website
has been online since 1994.

@@

As stopwords, we used the list that is provided by the Gensim Python
library\footnote{http://radimrehurek.com/gensim/} and we added following very
common words in the existing dataset: ``data'', ``based'', ``new'', ``project'',
``university'', ``student'', ``students'', ``research'', ``study'', ``program'',
``development'', ``study'', ``studies'', ``provide'', ``use''. We also removed 
the rare words that appear less than 5 times in the collection. 
Finally, we obtained a smaller corpus per framework (see table \ref{eutable}).

@@

The NSF organizes its research and education support through seven 
directorates, each encompassing several disciplines:

\begin{itemize}
\item Biological Sciences
\begin{description}
\item Molecular, Cellular and Organismal biology, Environmental science
\end{description}
\item Computer and Information Science and Engineering
\begin{description}
\item Fundamental computer science, Computer and networking systems, 
Artificial Intelligence
\end{description}
\item Engineering
\begin{description}
\item Bioengineering, Environmental systems, Civil and Mechanical systems, 
Chemical and Transport systems, Electrical and Communications systems, Design
 and Manufacturing
\end{description}
\item Geosciences
\begin{description}
\item Geological, Atmospheric and Ocean sciences
\end{description}
\item Mathematical and Physical Sciences
\begin{description}
\item Mathematics, Astronomy, Physics, Chemistry and Materials science
\end{description}
\item Social, Behavioral and Economic Sciences
\begin{description}
\item Neuroscience, Management, Psychology, Sociology, Anthropology, 
Linguistics, Science of science policy and economics
\end{description}
\item Education and Human Resources
\begin{description}
\item Science, Technology, Engineering and mathematics education at every level
\end{description}
\end{itemize}

The data about the awards
\footnote{https://www.nsf.gov/awardsearch/download.jsp} are open, free and can
be used in research and applications. The dataset that we chose for the
following analysis contains information about all the projects that got funded
by the NSF from 1994 to 2017.

In the provided dataset, the information is splitted into XML files, separated
per year, from 1994-2017. In order to keep the correlation to time and be able
to compare the two datasets later, we splitted the projects based on the year
they were initialized (see table \ref{usatable}).

@@

Just like the EU dataset, we followed the same preprocessing steps to obtain the
final corpus. As stopwords, we used the list that is provided by the Gensim
Python library and we removed following very common words in the existing
dataset: ``data'', ``work'', ``based'', ``new'', ``project'', ``university'', ``student'', ``students'', ``research'', ``study'', ``program'', ``development'', ``study'', ``studies'', ``provide'', ``use'', ``understanding'', ``important'', ``support'', ``proposed''. We also removed the rare words that appear less than 5 
times in the collection. Finally, we obtained a smaller corpus per framework 
(see Table \ref{usatable}).

@@

\subsection{Model inference}

For the implementation, we used the Gensim package (Radim Řehůřek,
2009)\cite{rehurek_lrec} which implements an efficient online LDA algorithm, to
extract the topics of a given corpus. Gensim uses a fast implementation of
online LDA parameter estimation based on \emph{Online Learning for Latent Dirichlet
Allocation}.\cite{onlineLDAvb} The algorithm runs in constant memory, which
means that the size of the training corpus does not affect memory footprint, so
it can process corpus larger than RAM.

@@

The implementation was run on a computer with an Intel® Xeon® Processor E5-1410
(10MB Cache, 2.8 GHz, 4 cores, 8 threads) and RAM=12 GB. Check the Appendix for 
the actual implementation.

@@

\todo{This is the heart of the thesis, but it is difficult to make
  sense of it. Make sure that when you mention a topic, the topic
  stands out of the text. It could be in italics, or some other font.
  Right now it is not easy to detect it in the discussion. From what I
  understand, the topics you mention are the top topics over all the
  years. You should then list them all in a table, in decreasing order
  of popularity. Also you should show the wordclouds. Right now there
  is only a reference to a figure at the appendix, which is not at the
  appendix at all.
}

In general, the most popular subject in European Union over the years
is Information systems. In our dataset we found projects
related to almost all kinds of information systems, and especially
Quality management systems (1994-2002), Management information
systems, Network information systems and Information systems as a
service to all kinds of organizations, including universities and
companies. Since technology has an impact in every aspect nowadays, it
is normal to have such popularity over the years and it's becoming
popular in education, as well. The same period of time, Innovation in
Business emerged as a trending topic, with projects related to new
technologies and software systems in the industry and how they assist
the businesses and the market. In addition, until 2002 a very popular
subject related to Information systems, was Applications in mobile
networks, embedded and broadband systems, which nowadays has lost part
of its popularity in research.

Another popular subject that appears the last 22 years is Materials research;
the projects related to material research are mostly related to the discovery
and design of new materials, with an emphasis on solids. Materials properties
and performance are used in many disciplines, like chemistry, physics and
engineering.

The third most popular subject is Molecular biology and Genetics.
These projects are related to the cellular molecules that carry out
the biological processes essential for the cell's functions and
maintenance. Especially from 1998 up to day, there is a focus in
Health care and Diagnostics, and for the last decade researchers are
studying the cancer detection and treatment, which is a common cause
of death in EU and the second most common cause of death in USA.
\todo{It seems now that you are discussing topics that are not popular
  throughout the period, but in the latest frameworks?}

Moreover, Environmental, Social and Economic Sustainability is a
subject very important to the research community and many researchers
work on this up to day. 
The European Union is considered by some to have the most extensive
environmental laws of any international organization, started by the
European Economic Community (EEC), which is the beginning of the EU's
environmental policy. Social and Economic policies are also popular in
the EU region over the last decade. Also, the last two decades there
is a focus on the Green energy production, like hydropower, solar and
nuclear energy.

During the 6th and the 7th framework programme, International Research and
conferences came up in our dataset as a trending topic. Researchers have a great
interest to collaborate with other colleagues from all over the world and share
their knowledge. A big part of the projects funded by the EU that is related to
international research, is also related to Physics, in general, and High energy
physics in EU (see Appendix: Wordclouds of topics of the EU).

\subsection{Discovering topics on the USA projects}

\todo{The same comments as in the previous subsection apply here. Make
  sure you explain how you picked the topics you discuss here. Make
  sure they stand out.}

With STEM being the top priority in USA, since this topic shows up in all the years
and according to the plot in figure \ref{usatrend} STEM is the topic with the highest proportion
of projects compared to the other topics (read section ), 
most of the projects founded by
NSF are related to engineering, computer science, mathematics and
information systems, in general. The subjects STEM \& Education,
Software engineering, Information systems and Mathematics, arise as
topics in all these years, from 1994-2017.

While NSF promotes the progress of science, it also enhances the efforts to
maintain the national health, prosperity, and welfare. Over the last 22 years,
many projects are related to the environment effects of climate change. Topics
like Environmental Science, Atmospheric \& Ocean sciences, Geology, Climate
change and its impact in the United States emerge during all these years. The
last three years there is a focus on the Green energy production, like
hydropower and wind power. Among with climate change, there is many research
related to Biology and Evolution. A less popular subject that appears during the
years 1994-1997 and 2007-2013 is Molecular, Cellular \& Organismal biology and
Genetics. These projects mostly refer to applied genetics, as well as cell
functionality. Topics include receptor biology, regulation of gene expression,
biomedical engineering, patterns of inheritance and molecular genetics.

Another popular subject that appears the last 22 years is Materials and
Chemistry; the projects related to material research are mostly refer to the
materials' properties and applications in engineering. The last 3 years,
researchers from the USA focus on materials applications and design in
mechanical engineering, bioengineering and chemical \& transport systems. In
2002-2006 many projects are related to the subject of transport that is
sustainable in the senses of social, environmental and climate impacts (Green
transportation).

From 1994-2006 and the last three years, International Research and
conferences came up as a trending topic. Researchers have a great 
interest to collaborate with other
colleagues from all over the world and share their knowledge. A big
part of the projects funded by the EU that is related to international
research, is also related to Physics, in general, and High energy
physics in EU (see Appendix: Wordclouds of topics of the EU).
\todo{Are you talking about the EU, or the US here?}

NSF shows a great interest in international activities across all NSF
supported disciplines. The goal of these projects is to support high
quality research and education, which could not occur without
international collaboration. NSF seeks to catalyze a higher level of
international engagement in the US science and engineering community
by promoting international conferences and university projects and
partnerships \todo{You have said the same in the previous paragraph.}.
An interesting fact regarding international partnerships is that the
European Commission and the Government of the United States of America
concluded on 17 October 2016 an Implementing Arrangement that
simplifies the cooperation between Horizon 2020 projects and US
entities. European and American researchers will be able to work
together more closely on projects funded under Horizon 2020, the
European Union's research and innovation
programme\footnote{\url{http://ec.europa.eu/research/iscp/index.cfm?pg=usa}}.

From 2002 up to day, Economic and Social Sciences emerge as a topic in our
dataset. Researchers show a great interest in identifying policies which help
the communities and they also do lots of Qualitative and Quantitative research
to investigate the size or extent of particular issues or trends in society.

Finally, a topic that used to be popular in the past, in 1994-1997 and
2007-2013, is High energy Physics in USA, which has lost its popularity
nowadays. The last 5 years, there is an interest around STEM and its
applications in manufacturing and engineering.

\subsection{Topic evolution over time}

>>>>

\usepackage[justification=centering]{caption}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{algorithm2e}
\usepackage{url}
\usepackage{subfig}
\usepackage{listings}
\usepackage[utf8]{inputenc}
\usepackage{color}
\usepackage{geometry} 
\usepackage{float}

@@

Today most research is funded by competitive open calls. 
The European Union countries are encouraged to
invest 3\% of their GDP in research and development by 2020. EU's
latest research programme, Horizon 2020, allows researchers to
identify new opportunities and directions in almost any field of
research. The programme's goals are to strengthen the EU's position in
science, support industrial innovation and address social concerns,
such as climate change, green transportation and energy.

@@

Latent Semantic Indexing was one of the earliest methods for finding
relationships between documents and the words that occur in
them\cite{Deerwester90indexingby}. In LSI a term-document matrix is
created by analyzing the corpus, where the rows correspond to words
and columns to documents. Each element in this sparse matrix describes
the number of times a word occurs in a document but this term count
can also be weighted with for instance TF-IDF\footnote{TF-IDF stands 
for term frequency - inverse document frequency, and the TF-IDF weight is 
a weight often used in information retrieval and text mining. This weight 
is a statistical measure used to evaluate how important a word is to 
a document in a collection or corpus. The importance increases
proportionally to the number of times a word appears in the document (TF) but
is offset by the frequency of the word in the corpus (IDF).}. Letting
each word represent a dimension in a very high dimensional space, a
document can be seen as a vector with components corresponding to its
weighted term counts\cite{Salton:1988:TAA:54259.54260}.

@@

\subsection{EU dataset}
\label{EUdataset}
CORDIS (Community Research and Development Information Service) is the European
Commission's primary public repository and portal to disseminate information on
all EU-funded research projects and their results in the broadest sense. The
website includes all public information held by the Commission (project
fact sheets, publishable reports and deliverables), editorial content to support
communication and exploitation (news, events, success stories, magazines) and
comprehensive links to external sources such as open access publications and
websites. CORDIS is managed by the Publications Office of the European Union, on
behalf of the European Commission's research Directorates-General and Agencies.
CORDIS content dates back to the origin of the service in 1990 and the website
has been online since 1994.

@@

As stopwords, we used the list that is provided by the Gensim Python
library\footnote{http://radimrehurek.com/gensim/} and we added following very
common words in the existing dataset: ``data'', ``based'', ``new'', ``project'',
``university'', ``student'', ``students'', ``research'', ``study'', ``program'',
``development'', ``study'', ``studies'', ``provide'', ``use''. We also removed 
the rare words that appear less than 5 times in the collection. 
Finally, we obtained a smaller corpus per framework (see table \ref{eutable}). 
For the actual implementation, see \emph{eu iterations.py} in the Appendix.

@@

NSF is divided into the following seven directorates that support science and
engineering research and education: Biological Sciences, Computer and
Information Science and Engineering, Engineering, Geosciences, Mathematical and
Physical Sciences, Social, Behavioral and Economic Sciences, and Education and
Human Resources. Each is headed by an assistant director and each is further
subdivided into divisions like materials research, ocean sciences and behavioral
and cognitive sciences. Within NSF's Office of the Director, the Office of
Integrative Activities also supports research and researchers. Other sections of
NSF are devoted to financial management, award processing and monitoring, legal
affairs, outreach and other functions \footnote{\url{https://www.nsf.gov/about/research_areas.jsp}}. The directorates and their divisions are listed below:
\begin{itemize}
\item Biological Sciences (BIO)
\begin{itemize}
\item[] Biological Infrastructure (DBI)
\item[] Environmental Biology (DEB)
\item[] Emerging Frontiers (EF)
\item[] Integrative Organismal Systems (IOS)
\item[] Molecular and Cellular Biosciences (MCB)
\end{itemize}
\item Computer and Information Science and Engineering (CISE)
\begin{itemize}
\item[] Office of Advanced Cyberinfrastructure (OAC)
\item[] Computing and Communication Foundations (CCF)
\item[] Computer and Network Systems (CNS)
\item[] Information and Intelligent Systems (IIS)
\end{itemize}
\item Education and Human Resources (EHR)
\begin{itemize}
\item[] Graduate Education (DGE)
\item[] Research on Learning in Formal and Informal Settings (DRL)
\item[] Undergraduate Education (DUE)
\item[] Human Resource Development (HRD)
\end{itemize}
\item Engineering (ENG)
\begin{itemize}
\item[] Chemical, Bioengineering, Environmental and Transport Systems (CBET)
\item[] Civil, Mechanical and Manufacturing Innovation (CMMI)
\item[] Electrical, Communications and Cyber Systems (ECCS)
\item[] Engineering Education and Centers (EEC)
\item[] Emerging Frontiers and Multidisciplinary Activities (EFMA)
\item[] Industrial Innovation and Partnerships (IIP)
\end{itemize}
\item Environmental Research and Education (ERE)
\item Geosciences (GEO)
\begin{itemize}
\item[] Atmospheric and Geospace Sciences (AGS)
\item[] Earth Sciences (EAR)
\item[] Ocean Sciences (OCE)
\item[] Office of Polar Programs (OPP)
\end{itemize}
\item Integrative Activities (OIA)
\item International Science and Engineering (OISE)
\item Mathematical and Physical Sciences (MPS)
\begin{itemize}
\item[] Astronomical Sciences (AST)
\item[] Chemistry (CHE)
\item[] Materials Research (DMR)
\item[] Mathematical Sciences (DMS)
\item[] Physics (PHY)
\item[] Office of Multidisciplinary Activities (OMA)
\end{itemize}
\item Social, Behavioral and Economic Sciences (SBE)
\begin{itemize}
\item[] Behavioral and Cognitive Sciences (BCS)
\item[] National Center for Science and Engineering Statistics (NCSES)
\item[] Social and Economic Sciences (SES)
\item[] SBE Office of Multidisciplinary Activities (SMA)
\end{itemize}
\end{itemize}
The data about the awards
\footnote{https://www.nsf.gov/awardsearch/download.jsp} are open, free and can
be used in research and applications. The dataset that we chose for the
following analysis contains information about all the projects that got funded
by the NSF from 1994 to 2017.

In the provided dataset, the information is split into XML files, separated
per year, from 1994-2017. In order to keep the correlation to time and be able
to compare the two datasets later, we split the projects based on the year
they were initialized (see table \ref{usatable}).

@@

Just like the EU dataset, we followed the same preprocessing steps to obtain the
final corpus. As stopwords, we used the list that is provided by the Gensim
Python library and we removed following very common words in the existing
dataset: ``data'', ``work'', ``based'', ``new'', ``project'', ``university'', ``student'', ``students'', ``research'', ``study'', ``program'', ``development'', ``study'', ``studies'', ``provide'', ``use'', ``understanding'', ``important'', ``support'', ``proposed''. We also removed the rare words that appear less than 5 
times in the collection. Finally, we obtained a smaller corpus per framework 
(see Table \ref{usatable}). 
For the actual implementation, see \emph{usa iterations.py} in the Appendix.

@@

\subsection{Model inference}
\label{modelinference}
For the implementation, we used the Gensim package (Radim Řehůřek,
2009)\cite{rehurek_lrec} which implements an efficient online LDA algorithm, to
extract the topics of a given corpus. Gensim uses a fast implementation of
online LDA parameter estimation based on \emph{Online Learning for Latent Dirichlet
Allocation}.\cite{onlineLDAvb} The algorithm runs in constant memory, which
means that the size of the training corpus does not affect memory footprint, so
it can process corpus larger than RAM.

@@

The implementation was run on a computer with an Intel® Xeon® Processor E5-1410
(10MB Cache, 2.8 GHz, 4 cores, 8 threads) and RAM=12 GB. Check the Appendix for 
the actual implementation.
\pagebreak

@@

In figures \ref{wordcloudEU4} - \ref{wordcloudEU20} we show the 10 topics from FP4 to H2020 of the EU dataset, described by
the 20 top words with the highest posterior probability of belonging
to the topic. The size of each word is in proportion to each
probability; the bigger the word, the higher the probability. 
Each wordcloud describes a topic, which we can find intuitively (the label 
of each wordcloud corresponds to the title of the topic).

By comparing the thematic areas suggested by the EU
\footnote{\url{https://europa.eu/european-union/topics_en}} and the topics that
we found with our implementation, one can easily observe that the majority of
these topics are included in the wider categories of ``Science \& Technology''
and its subcategory ``Research and Innovation''. For instance, the topics
\emph{Molecular Biology} and \emph{Genetics} belong to the category ``Science \&
Technology'', \emph{Cancer detection \& treatment} to the category ``Health''
and ``Science'', \emph{Mathematics}, \emph{High energy physics}, \emph{Materials
\& Chemistry} and \emph{Information Systems} belong to ``Science \&
Technology''. Even though the topics that we listed above stand out as distinct
thematic areas, they are not listed in the EU's website.

The topics that we are going to discuss in this section are the ones that
stand out in most of the Framework Programmes.
In general, the most popular subject in European Union over the years
is \emph{Information systems}. In our dataset we found projects
related to almost all kinds of information systems, and especially 
\emph{Management information
systems}, \emph{Network information systems} and \emph{Information systems as a
service} to all kinds of organizations, including universities and
companies. Since technology has an impact in every aspect nowadays, it
is normal to have such popularity over the years and it's becoming
popular in education, as well. The same period of time, \emph{Innovation in
Business} emerged as a trending topic, with projects related to new
technologies and software systems in the industry and how they assist
the businesses and the market. In addition, until 2002 a very popular
subject related to \emph{Information systems}, was \emph{Applications in mobile
networks}, \emph{embedded and broadband systems}, which nowadays has lost part
of its popularity in research.

Another popular subject that appears the last 22 years is \emph{Materials research};
the projects related to material research are mostly related to the discovery
and design of new materials, with an emphasis on solids. Materials properties
and performance are used in many disciplines, like chemistry, physics and
engineering.

The third most popular subject is \emph{Molecular biology \& Genetics}.
These projects are related to the cellular molecules that carry out
the biological processes essential for the cell's functions and
maintenance. Especially from 1998 up to day, there is a focus in
Health care and Diagnostics, and for the last decade researchers are
studying the \emph{Cancer detection \& treatment}, which is a common cause
of death in EU and the second most common cause of death in USA.

Moreover, \emph{Environmental, Social and Economic Sustainability} is a
subject very important to the research community and many researchers
work on this up to day. 
The European Union is considered by some to have the most extensive
environmental laws of any international organization, started by the
European Economic Community (EEC), which is the beginning of the EU's
environmental policy. \emph{Social and Economic policies} are also popular in
the EU region over the last decade. Also, the last two decades there
is a focus on the \emph{Green energy production}, like hydropower, solar and
nuclear energy.

During the 6th and the 7th framework programme, \emph{International Research \& 
conferences} came up in our dataset as a trending topic. Researchers have a great
interest to collaborate with other colleagues from all over the world and share
their knowledge. A big part of the projects funded by the EU that is related to
\emph{international research}, is also related to \emph{Physics}, in general, 
and \emph{High energy physics}.

\begin{figure}[H]
\begin{tabular}{rcl}
\subfloat[Management Systems \& Software engineering]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic0_7000.png}} &
\subfloat[Agriculture, Fisheries and Food]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic1_7000.png}}&
\subfloat[Molecular biology, Genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic2_7000.png}}\\
\subfloat[EU transport]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic3_7000.png}}&
\subfloat[Digital systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic4_7000.png}} &
\subfloat[Materials processing]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic5_7000.png}}\\
\subfloat[Genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic6_7000.png}} &
\subfloat[Policy for environmental sustainability]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic7_7000.png}}&
\subfloat[Energy technologies]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic8_7000.png}}\\
\subfloat[Applications in mobile networks]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP4 (1994-1998)}
\label{wordcloudEU4}
\end{figure}

\begin{figure}[H]
\begin{tabular}{rcl}
\subfloat[Management software]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic0_7000.png}} &
\subfloat[Hydropower / Energy production]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic1_7000.png}}&
\subfloat[Information technologies]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic2_7000.png}}\\
\subfloat[Quality control systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic3_7000.png}}&
\subfloat[Health care / Medical genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic4_7000.png}} &
\subfloat[Environmental sustainability and policies]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic5_7000.png}}\\
\subfloat[Molecular biology (genes)]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic6_7000.png}} &
\subfloat[Optical quantum computing]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic7_7000.png}}&
\subfloat[Education in Nuclear physics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic8_7000.png}}\\
\subfloat[Education, training]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP5 (1998-2002)}
\label{wordcloudEU5}
\end{figure}

\begin{figure}[H]
\begin{tabular}{rcl}
\subfloat[International research]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic0_7000.png}} &
\subfloat[Biology]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic1_7000.png}}&
\subfloat[Embedded systems design]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic2_7000.png}}\\
\subfloat[Molecular biology, genes]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic3_7000.png}}&
\subfloat[Evolution / Formation]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic4_7000.png}} &
\subfloat[Sustainable production / Energy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic5_7000.png}}\\
\subfloat[Economic \& Social policy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic6_7000.png}} &
\subfloat[Environmental management services]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic7_7000.png}}&
\subfloat[Materials]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic8_7000.png}}\\
\subfloat[Information systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP6 (2002-2006)}
\label{wordcloudEU6}
\end{figure}

\begin{figure}[H]
\begin{tabular}{rcl}
\subfloat[Cancer detection \& treatment]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic0_7000.png}} &
\subfloat[(High energy) physics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic1_7000.png}}&
\subfloat[Energy production]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic2_7000.png}}\\
\subfloat[Social Sciences]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic3_7000.png}}&
\subfloat[Molecular Biology / Genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic4_7000.png}} &
\subfloat[Information systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic5_7000.png}}\\
\subfloat[International research]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic6_7000.png}} &
\subfloat[Mathematics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic7_7000.png}}&
\subfloat[Materials]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic8_7000.png}}\\
\subfloat[Molecular biology]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP7_wordclouds/EUFP7_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP7 (2007-2013)}
\label{wordcloudEU7}
\end{figure}

\begin{figure}[H]
\begin{tabular}{rcl}
\subfloat[Molecular biology]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic0_20000.png}} &
\subfloat[Environmental sustainability \& food production]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic1_20000.png}}&
\subfloat[Information systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic2_20000.png}}\\
\subfloat[Training]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic3_20000.png}}&
\subfloat[Materials]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic4_20000.png}} &
\subfloat[Environmental sustainability / Green energy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic5_20000.png}}\\
\subfloat[Physics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic6_20000.png}} &
\subfloat[Business \& Innovation]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic7_20000.png}}&
\subfloat[Cancer detection \& treatment]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic8_20000.png}}\\
\subfloat[Economic \& Social policy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_H2020_wordclouds/EUH2020_topic9_20000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU H2020 (2014-2020)}
\label{wordcloudEU20}
\end{figure}

\vfill
\subsection{Discovering topics on the USA projects}

In figures \ref{wordcloudUSA4} - \ref{wordcloudUSA20} we show the 10 topics 
from 1994 to 2017 of the USA dataset, described by
the 20 top words with the highest posterior probability of belonging
to the topic.
By comparing the the areas of interest in the US, through its directorates, 
and our topics, we see that all the topics that we found correspond to either 
a directorate or one of the divisions. Below, we discuss the topics that
stand out in most of the years.

With \emph{STEM} being the top priority in USA, since this topic shows up in all the years
and according to the plot in figure \ref{usatrend} \emph{STEM} is the topic with the highest proportion
of projects compared to the other topics (read section \ref{usaevolution}), 
most of the projects founded by
NSF are related to engineering, computer science, mathematics and
information systems, in general. The subjects \emph{STEM \& Education},
\emph{Software engineering}, \emph{Information systems} and \emph{Mathematics}, arise as
topics in all these years, from 1994-2017.

While NSF promotes the progress of science, it also enhances the efforts to
maintain the national health, prosperity, and welfare. Over the last 22 years,
many projects are related to the environment effects of climate change. Topics
like \emph{Environmental Science, Atmospheric \& Ocean sciences, Geology, Climate
change} and its impact in the United States emerge during all these years. The
last three years there is a focus on the \emph{Green energy production}, like
hydropower and wind power. Among with climate change, there is many research
related to \emph{Biology and Evolution}. A less popular subject that appears during the
years 1994-1997 and 2007-2013 is \emph{Molecular, Cellular \& Organismal biology and
Genetics}. These projects mostly refer to applied genetics, as well as cell
functionality. Topics include receptor biology, regulation of gene expression,
biomedical engineering, patterns of inheritance and molecular genetics.

Another popular subject that appears the last 22 years is \emph{Materials and
Chemistry}; the projects related to material research are mostly refer to the
materials' properties and applications in engineering. The last 3 years,
researchers from the USA focus on materials applications and design in
mechanical engineering, bioengineering and chemical \& transport systems. In
2002-2006 many projects are related to the subject of \emph{Transport} that is
sustainable in the senses of social, environmental and climate impacts (\emph{Green
transportation}).

NSF shows a great interest in \emph{International activities} across all NSF
supported disciplines. The goal of these projects is to support high
quality research and education, which could not occur without
international collaboration.
An interesting fact regarding international partnerships is that the
European Commission and the Government of the United States of America
concluded on 17 October 2016 an Implementing Arrangement that
simplifies the cooperation between Horizon 2020 projects and US
entities. European and American researchers will be able to work
together more closely on projects funded under Horizon 2020, the
European Union's \emph{Research and Innovation}
programme\footnote{\url{http://ec.europa.eu/research/iscp/index.cfm?pg=usa}}.

From 2002 up to day, \emph{Economic and Social Sciences} emerge as a topic in our
dataset. Researchers show a great interest in identifying policies which help
the communities and they also do lots of Qualitative and Quantitative research
to investigate the size or extent of particular issues or trends in society.

Finally, a topic that used to be popular in the past, in 1994-1997 and
2007-2013, is \emph{High energy Physics} in USA, which has lost its popularity
nowadays. The last 5 years, there is an interest around STEM and its
applications in \emph{manufacturing and engineering}.
